Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-04-02 WSSSPE2_report/WSSSPE2_review.tex
Sections 4 and 5: format change.
putting tables back
adding stuff about lightning talks in abstract, section 4, conclusions fixing some affils in appendix adding note in 3.2

\begin{abstract}      
This technical report records and discusses the Second Workshop on Sustainable
Software for Science: Practice and Experiences (WSSSPE2). The workshop used an
alternative submission and peer-review process, which led to a set of papers
divided across five topic areas: exploring sustainability; software development
experiences; credit \& incentives; reproducibility \& reuse \& sharing; and code
testing \& code review. The report includes a description of the submission and
review process, the workshop keynote presentations, a discussion on
sustainability, and the five discussions from the topic areas. The report also
contains a description of potential actions that were proposed in the the five
discussions.
\end{abstract}

@@

%\url{http://dx.doi.org/10.6084/m9.ﬁgshare.1243288}

Neil Chue Hong is currently serving as the Director of the Software
Sustainability Institute (SSI), United Kingdom. Chue Hong started with making
the point that world-leading research relies on software, quoting Kersten Kleese
Van Dam of the Pacific Northwestern National Laboratory via the campaign
\href{http://change.org}{change.org}, ``\emph{Today there are very few science
areas left which do not rely on IT and thus software for the majority of the
their research work. More importantly key scientific advances in experimental
and observational science would have been impossible without better software.}''
He also cited Daniel Katz, Software Infrastructure for Sustained Innovation
(SI2) Program Director of the National Science Foundation, ``\emph{Scientific
discovery and innovation are advancing along fundamentally new pathways opened
by development of increasingly sophisticated software. Software is an integral
enabler of computation, experiment and theory, and directly responsible for
increased scientific productivity and enhancement of researchers'
capabilities.}''

@@

\note{Lead: Choi. Volunteers welcome.
\href{http://wssspe.researchcomputing.org.uk/wssspe2/lightning-talks/}{Slides}}

Lightning talks is a new feature in WSSSPE2. Every talk in it were given only
two minutes and the space of one slide. A total of 18 lightning talks were
selected in reverse alphabetic order of speakers' last names.
\begin{enumerate}
\item Greg Wilson: Close Enough for Scientific Work

@@

\katznote{yes, along with a change to breakout groups that were intended to
define actions, then potentially become longer lasting working groups to perform
those actions}

\katznote{yes, please feel free to add something in the abstract and intro}

@@

\begin{table*}[th]
 \centering%\begin{center}
\caption{Enablers of software sustainability, with 0 to 10 `*'s roughly
indicating the fraction of attendees who voted for an item as important.}
\label{tb:software_sustainability_enablers}
    %\begin{scriptsize}
    %\begin{tabular}{ | p{1.65cm} | p{10.0cm} |}
     \begin{tabular}{ | p{1.9cm} | p{11.5cm} |}
    \hline
  Importance & Item \\ \hline \hline
********** & healthy and vibrant communities; vibrant community to champion software \\ \hline
********** & designing for growth and extension -- open development \\ \hline
******* & culture in community for reuse \\ \hline
**** & portability \\ \hline
**** & culture in dev community to support transition between developers \\ \hline
*** & interdisciplinary people: science + IT experience \\ \hline
** & planning for end of life \\ \hline
** & make smart choices about dependencies \\ \hline
* & thinking of software as product lines -- long term vs.~short term view \\ \hline
 & not all communities need new software \\ \hline
 & converting use into resources \\ \hline
    \end{tabular}
    %\end{scriptsize}
%\end{center}   
\end{table*}

\begin{table}[th]
\caption{Barriers to software sustainability, with 0 to 10 `*'s roughly
indicating the fraction of attendees who voted for an item as important.}  
\label{tb:software_sustainability_barriers} 
 \centering
  %\begin{scriptsize}
  %\begin{tabular}{ | p{1.65cm} | p{10.0cm} |}
 \begin{tabular}{ | p{1.9cm} | p{11.5cm} |}
   \hline
  Importance & Item \\ \hline \hline
******* & lack of incentives, including promotion and tenure process; promotion
and tenure process in academic is incompatible with sustainability \\ \hline
*****  & absent or poor documentation \\ \hline
***** & funding to ensure sustainability is difficult to obtain \\ \hline
**** & developers are not computer scientists; don't have software engineering
practices (in particular, those needed to scale-up projects to support and be
developed by a large sustainable community) \\ \hline
*** & overreliance on one or two people -- the `bus test'\footnote{``Bus test'' in Table~\ref{tb:software_sustainability_barriers} refers to the smallest number of key people a project would need to lose to become non-viable (the larger the number, the healthier the project).} \\ \hline
** & rate of change of underlying technologies \\ \hline
** & lack of business models for sustainability \\ \hline
* & lack of training for how to build sustainability into the system \\ \hline
 & maintenance for software is not visible, appears to ``just happen'' \\ \hline
 & licensing issues \\ \hline
 & staff turnover -- lack of continuity \\ \hline
\end{tabular}
% \end{scriptsize}
\end{table}

@@

@@

\\ Gabrielle Allen       &  U Illinois

@@

\\ Wolfgang Bangerth & Texas A\&M

@@

\\ Carl Boettiger        &  UCSC/ropensci

@@

\\ Steven R.~Brandt      &  Louisiana State University

@@

\\ John W.~Cobb

@@

\\ Marcus D.~Hanwell     &  Kitware

@@

\\ Daniel S.~Katz        &  U Chicago \& Argonne

@@

\\ Hilmar Lapp           &  National Evolutionary Synthesis Center (NESCent) and Duke University

@@

\\ Richard Littauer      &  MIT/University of Saarland

@@

\\ Robert H.~McDonald    &  Indiana University - Data to Insight Center/Libraries

@@

\\ Cody J.~Permann       &  Idaho National Laboratory
\\ John W.~Peterson      &  Idaho National Laboratory

@@

\\ Ray Plante            &  NCSA/University of Illinois Urbana-Champaign

@@

\\ Karthik Ram           &  University of California, Berkeley. Berkeley Institute for Data Science and rOpenSci
\\ Mike Rilee            &  NASA/GSFC, Rilee Systems Technologies

@@

\\ Matt Turk             &  NCSA/UIUC
\\ Colin C.~Venters      &  University of Huddersfield

@@

\\ Nancy Wilkins-Diehr   &  SDSC

>>>>

\begin{abstract}      
This technical report records and discusses the Second Workshop on Sustainable
Software for Science: Practice and Experiences (WSSSPE2). The workshop used an
alternative submission and peer-review process, which led to a set of papers
divided across five topic areas: exploring sustainability; software development
experiences; credit \& incentives; reproducibility \& reuse \& sharing; and code
testing \& code review. The report includes a description of the submission and
review process, the workshop keynote presentations, a set of lightning talks, a discussion on
sustainability, and the five discussions from the topic areas. The report also
contains a description of potential actions that were proposed in the the five
discussions.
\end{abstract}

@@

%\url{http://dx.doi.org/10.6084/m9.figshare.1243288}

Neil Chue Hong is currently serving as the Director of the Software
Sustainability Institute (SSI), United Kingdom. Chue Hong started with making
the point that world-leading research relies on software, quoting Kersten Kleese
Van Dam of the Pacific Northwestern National Laboratory via a petition campaign at \katznote{add cite for the petition}
change.org, ``\emph{Today there are very few science
areas left which do not rely on IT and thus software for the majority of the
their research work. More importantly key scientific advances in experimental
and observational science would have been impossible without better software.}''
He also cited Daniel Katz, Software Infrastructure for Sustained Innovation
(SI2) Program Director of the National Science Foundation, ``\emph{Scientific
discovery and innovation are advancing along fundamentally new pathways opened
by development of increasingly sophisticated software. Software is an integral
enabler of computation, experiment and theory, and directly responsible for
increased scientific productivity and enhancement of researchers'
capabilities.}''

@@

\note{Lead: Choi. Volunteers welcome.
\href{http://wssspe.researchcomputing.org.uk/wssspe2/lightning-talks/}{Slides}}
 \todo{in the abstract and intro}

Lightning talks were a new feature in WSSSPE2. Since the workshop program was
mostly dedicated to discussions, the organizers wanted to give the attendees a
chance to also `make a pitch' for an idea, either representing a contributed paper
or something different.  18 attendees volunteered to participate in the lightning
talks, each given two minutes to talk, and one slide if they wanted to use one.
The talked were presented in the reverse alphabetic order of speakers' last names.

\begin{enumerate}
\item Greg Wilson: Close Enough for Scientific Work

@@

@@

\begin{table*}[ht]
 \centering
\caption{Enablers of software sustainability, with 0 to 10 `*'s roughly
indicating the fraction of attendees who voted for an item as important.}
\label{tb:software_sustainability_enablers}
  \begin{scriptsize}
  \begin{tabular}{ | p{1.65cm} | p{10.0cm} |}
   %  \begin{tabular}{ | p{1.9cm} | p{11.5cm} |}
    \hline
  Importance & Item \\ \hline \hline
********** & healthy and vibrant communities; vibrant community to champion software \\ \hline
********** & designing for growth and extension -- open development \\ \hline
******* & culture in community for reuse \\ \hline
**** & portability \\ \hline
**** & culture in dev community to support transition between developers \\ \hline
*** & interdisciplinary people: science + IT experience \\ \hline
** & planning for end of life \\ \hline
** & make smart choices about dependencies \\ \hline
* & thinking of software as product lines -- long term vs.~short term view \\ \hline
 & not all communities need new software \\ \hline
 & converting use into resources \\ \hline
    \end{tabular}
    \end{scriptsize}
\end{table*}

\begin{table}[ht]
\caption{Barriers to software sustainability, with 0 to 10 `*'s roughly
indicating the fraction of attendees who voted for an item as important.}  
\label{tb:software_sustainability_barriers} 
 \centering
  \begin{scriptsize}
  \begin{tabular}{ | p{1.65cm} | p{10.0cm} |}
% \begin{tabular}{ | p{1.9cm} | p{11.5cm} |}
   \hline
  Importance & Item \\ \hline \hline
******* & lack of incentives, including promotion and tenure process; promotion
and tenure process in academic is incompatible with sustainability \\ \hline
*****  & absent or poor documentation \\ \hline
***** & funding to ensure sustainability is difficult to obtain \\ \hline
**** & developers are not computer scientists; don't have software engineering
practices (in particular, those needed to scale-up projects to support and be
developed by a large sustainable community) \\ \hline
*** & overreliance on one or two people -- the `bus test'\footnote{``Bus test'' in Table~\ref{tb:software_sustainability_barriers} refers to the smallest number of key people a project would need to lose to become non-viable (the larger the number, the healthier the project).} \\ \hline
** & rate of change of underlying technologies \\ \hline
** & lack of business models for sustainability \\ \hline
* & lack of training for how to build sustainability into the system \\ \hline
 & maintenance for software is not visible, appears to ``just happen'' \\ \hline
 & licensing issues \\ \hline
 & staff turnover -- lack of continuity \\ \hline
\end{tabular}
 \end{scriptsize}
\end{table}

@@

\katznote{add discussion of keynotes, lightning talks, charge to working groups about building plans, etc.}

@@

\\ Gabrielle Allen       & University of Illinois Urbana-Champaign

@@

\\ Wolfgang Bangerth & Texas A\&M University

@@

\\ Carl Boettiger        &  University of California Santa Cruz \& rOpenSci

@@

\\ Steven R. Brandt      &  Louisiana State University

@@

\\ John W. Cobb

@@

\\ Marcus D. Hanwell     &  Kitware

@@

\\ Daniel S. Katz        &  University of Chicago \& Argonne National Laboratory

@@

\\ Hilmar Lapp           &  National Evolutionary Synthesis Center (NESCent) \& Duke University

@@

\\ Richard Littauer      &  MIT \& University of Saarland

@@

\\ Robert H. McDonald    &  Data to Insight Center/Libraries, Indiana University

@@

\\ Cody J. Permann       &  Idaho National Laboratory
\\ John W. Peterson      &  Idaho National Laboratory

@@

\\ Ray Plante            &  NCSA, University of Illinois Urbana-Champaign

@@

\\ Karthik Ram           &  Berkeley Institute for Data Science, University of California Berkeley \& rOpenSci
\\ Mike Rilee            &  NASA/GSFC \& Rilee Systems Technologies

@@

\\ Matt Turk             &  NCSA, University of Illinois Urbana-Champaign
\\ Colin C. Venters      &  University of Huddersfield

@@

\\ Nancy Wilkins-Diehr   &  San Diego Supercomputer Center, University of California San Diego
